@c -*-texinfo-*-

@c  Copyright (c) 2023 Blake McBride
@c  All rights reserved.
@c
@c  Redistribution and use in source and binary forms, with or without
@c  modification, are permitted provided that the following conditions are
@c  met:
@c
@c  1. Redistributions of source code must retain the above copyright
@c  notice, this list of conditions and the following disclaimer.
@c
@c  2. Redistributions in binary form must reproduce the above copyright
@c  notice, this list of conditions and the following disclaimer in the
@c  documentation and/or other materials provided with the distribution.
@c
@c  THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
@c  "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
@c  LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
@c  A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
@c  HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
@c  SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
@c  LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
@c  DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
@c  THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
@c  (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
@c  OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.


@node Desktop Application

@html
@include style.css
@end html

@chapter Desktop Applications

Portable desktop
applications can be built utilizing @emph{Kiss} in conjunction with
Electron (@uref{https://electronjs.org}).  These
applications would be portable to Linux, macOS, Windows, BSD, and other
platforms with a single codebase.

Both the front-end and back-end are built exactly as a browser-based application would be built.
The only difference is how the system is run.

@section Setup

Before a desktop application can be built, the @emph{electron} system must be installed.
This, in turn, is dependent on @emph{nodejs} and @emph{npm}.

@subsection Installing Nodejs & Npm

Full instructions to install @emph{nodejs} and @emph{npm} can be found at
@uref{https://nodejs.org}

Linux systems that use the @emph{apt} package manager can install  @emph{nodejs} and @emph{npm} with the following command: 

@example
sudo apt install nodejs npm
@end example

Linux systems that use the @emph{dnf} package manager can install  @emph{nodejs} and @emph{npm} with the following command: 

@example
sudo dnf install nodejs
@end example

Other systems can install  @emph{nodejs} and @emph{npm} as instructed at
@uref{https://nodejs.org}

@subsection Installing Electron

You can install @emph{electron} on any OS from within the @emph{KISS} application with the following command:

@example
npm install electron --save-dev
@end example

@section Building & Running

The @emph{KISS} application is built no differently than if it were a browser application.  You can use the following command:

You can build the system with the following command:

@example
./bld build              [all OS except Windows]
    or
bld build                [Windows]
@end example

The system can be run as an application by executing the following command:

@example
npm start
@end example

@section Debug Mode

If you are developing and need the front-end developer tools, simply edit the file named @emph{start-electron.js}
Change the value of the variable named @emph{devMode} from @emph{false} to @emph{true}.  Then just start normally.

When you do this, you will get the beloved developer tools window along with the application.
