@c -*-texinfo-*-

@c  Copyright (c) 2026 Blake McBride
@c  All rights reserved.
@c
@c  Redistribution and use in source and binary forms, with or without
@c  modification, are permitted provided that the following conditions are
@c  met:
@c
@c  1. Redistributions of source code must retain the above copyright
@c  notice, this list of conditions and the following disclaimer.
@c
@c  2. Redistributions in binary form must reproduce the above copyright
@c  notice, this list of conditions and the following disclaimer in the
@c  documentation and/or other materials provided with the distribution.
@c
@c  THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
@c  "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
@c  LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
@c  A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
@c  HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
@c  SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
@c  LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
@c  DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
@c  THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
@c  (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
@c  OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.

@node External Integrations

@html
@include style.css
@end html

@chapter External Integrations

@emph{Kiss} ships with built-in support for connecting your application
to systems beyond itself --- both calling external services and being
called by them --- as well as integrating with modern AI tooling.  This
chapter notes that these facilities exist; full details, including
class names, configuration, and worked examples, are in the @emph{Kiss}
book.

@section Integrating With External Systems

@itemize @bullet
@item @emph{REST Client} --- make HTTP calls out to third-party APIs.
@item @emph{REST Server} --- accept inbound HTTP calls from external
systems and webhooks for clients that do not speak the @emph{Kiss}
JSON-RPC protocol.
@end itemize

@section LLM Integration

@itemize @bullet
@item @emph{LLM Providers} --- built-in support for OpenAI, Anthropic,
and Ollama.
@item @emph{MCP Server} --- expose application tools to AI assistants
via the Model Context Protocol.
@item @emph{Embeddings Database} --- vector storage for
retrieval-augmented generation (RAG).
@end itemize

@section Security

@itemize @bullet
@item @emph{OAuth 2.1 Resource Server} --- validate bearer tokens
issued either by this same Kiss app acting as the authorization server
(see the next bullet) or by an external provider such as Auth0, Okta,
Keycloak, Microsoft Entra, or Google.  When configured,
@code{MCPServerBase} subclasses automatically require valid tokens; no
code change is needed.  Configuration lives in
@code{application.ini}; the discovery document is published at
@code{/.well-known/oauth-protected-resource}.
@item @emph{OAuth 2.1 Authorization Server} --- issue tokens directly
from Kiss: authorization-code flow with mandatory PKCE, token endpoint
with refresh-token rotation, dynamic client registration (RFC 7591),
and the RFC 8414 metadata and JWKS endpoints.  Persists clients,
refresh tokens, and signing keys to a private SQLite database that is
independent of the application's main database --- no shared schema,
no operator setup beyond pointing at a file path.  Enabled by a single
@code{application.ini} key.
@item @emph{OAuth 2.1 Client} --- obtain tokens from a @emph{remote}
OAuth 2.1 server and present them on outbound calls.  Kiss drives the
authorization-code + PKCE flow with automatic discovery (RFC 9728 ->
RFC 8414 / OpenID Connect) and dynamic client registration when no
@code{ClientId} is configured, captures the redirect at a built-in
callback servlet, and refreshes access tokens automatically.  Multiple
remote providers are supported, each declared in its own
@code{[OAuthClient @var{name}]} section of @code{application.ini}.
Cached registrations and tokens are stored in the same @code{oauth.db}
the authorization server uses.
@end itemize

Each of the three roles is inert until configured and runs
independently or together.  See @code{OAuth.md} at the repository root
for the full how-to on all three, and the @emph{Kiss} book for the
in-depth treatment.
